2nd_draft

### 2020-03-19 src/2nd_draft.Rmd
added text from first draft, strat revising..

1. data processing
```{r, warning=F, message=F}
## Aggregated Data
assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))[,-1]
#assessment_2020 <- read.csv(here("data","assessment_2020.csv"))[,-1]
#assessment_2020.full <- read_excel(here("data/Assessment", "2020 Only.xlsx"))
```

@@

>>>>

## Summary

Our client is a real estate property tax analyst who analyzes the market trend and makes market predictions for his company and future clients. The main objective of our project is to accurately predict future mill rate (property tax) in metro Vancouver for the following 3 property tax classes: Tax class 1: Residential, Tax class 5: Light industry and Tax class 6: Business and other. Data cleaning, exploratory data analysis are used in this project to analyze the relationship between mill rate and other factors. Data cleaning is performed to aggregate our data into summary statistics. Exploratory analysis has shown there is a linear trend in mill rates over the years; it has also shown there is a fairly strong correlation between mill rate and average assessment per property in different municipalities. Ordinary Linear Model and a variety of its variants are used to predict the mill rate. A full assessment of the performance of these models is shown below.

## Introduction

Our goal is to predict mill rates in 2020. We also seek to identify which explanatory variables are the most important in determining mill rates. Every year, the assessment value of each property is released at the beginning of the year, however, the mill rate is still unknown until Spring. Prediction of mill rate is a focus of interest because it gives an approximate property tax to pay for property owners. It is also important because it might affect future buyers’ purchasing incentives. The property tax rate has a fairly small margin to change. Mill rate is adjusted based on the total assessment in each city so the municipal government can use tax earning (total assessment * mill rate) to match their annual expense in order to balance the city’s budget.

Correlation between the mill rate and each explanatory variable will be asses to pick the essential variables in our model. Then, a variety of linear models will be fitted using our selected variable. The best model is selected based on its prediction power and goodness of fit.

## Data Description

For exploratory data analysis and building predictive model, we used past property assessment data, which is provided by our client. It contains information about some commercial and residential properties in all cities in Metro Vancouver from 2016 to 2019. We selected variables that we believe were relevant to the property tax prediction. These selected variables are:

- Mill Rates (2016-2019)
- Assessment (2016-2019)
- Tax Class Code (01, 05, 06)
- Area Code (Municipality)
- Year
- Land Total (2016 - 2019)
- Improvement Total (2016 - 2019)
- Number of properties

As the client suggested, we have also collected external data from the Government of British Columbia which could be associated with the tax rate:

- Municipal Budget of Cities in Metro Vancouver (Taxes Imposed & Collected, Schedule 706)

We decided to calculate the sum of TotalAssessedValue, LandAssessedValue, ImprovementAssessedValue, and Budget for each combination of municipality and the tax class to reduce the dimension of our data. The mill rate is unique for each combination of municipality and tax class, so we have added this information into our data for each municipality and its tax classes.

## Methods

### Exploratory Analysis

Before we made any prediction on the future mill rate of Metro Vancouver’s real estate market, we did some exploratory analysis to explore and visualize the main characteristics of our dataset and found relationships between Mill Rate vs. Assessment, Mill Rate vs. Municipal Budget, Mill Rate vs. Land Total, Mill Rate vs. Imrpovement Total and Mill Rate vs. Number of Properties.

The municipal government wants their tax income to match their annual spending in order to balance their bedget. Given this piece of information, we want to find and model the relationship between municipal tax income and municipal budget. We plotted a graph of tax income for three different tax classes (01, 05, and 06) vs. budget through time to see the trend of tax income in different municipalities to make sure our assumption about the data is correct. We chose to illustrate such relationship using scatter plot below for Vancouver, Richmond and Burnaby as they are the most representing city in Metro Vancouver.

Mill rate is mainly affected by assessment and municipal budget, so we have plotted scatter plot of mill rate vs. assessment and mil rate vs. municipal budget to see the correlation between each pair of the two factors. ANOVA analysis is also performed to see the correlation between mill rate vs. tax class, mill rate vs. year and mill rate vs. municipality.

### Measure of goodness of fit and prediction power

In this study, we built a few linear models and evaluated their performances by goodness of fit and prediction power, that is, how well the model explains the data and how well it can predict future values. The definition of goodness of fit and prediction power is given below.

 - **Goodness of fit** is defined as the extent to which the sample data are consistent with the model, which examines how well the model explains the data. R squared and adjusted R squared are the most well known measures of goodness of fit and higher R squared and adjusted R squares indicate better goodness of fit.

 - **Prediction power** measures how well models can predict the future values. We use mean squared prediction power to compare the prediction performance across all of our fitted linear predictive models. To do that, we will divide the data set into training data - used to build our models, and testing data - used to evaluate the prediction power of our models.

### Linear model

For our linear model, Year, TaxClassCode, Municipalities, Assess Total, Land Assessment Total, Improvement Assessment total, Number of properties and tax (budget) are considered. We first explore the full linear model using all the available features. From our exploratory data analysis, we have found Vancouver, Richmond, Surrey and Burnaby are outliers. Further prediction model will be built based on these findings. We are also intesterd in how different tax class can affect our predictive model, so models based on different tax classes will also be considered in the future. 
For our reduced model, we build our linear predictor using only the significant variables from the full model (See table 4). The reduced model's performence in terms of goodness of fit will decrease with fewer features, however, the performance of prediction power might be improved.

### Ridge, Lasso and Elastic Net

Other than simple linear regression, we are also interested in the performance of Ridge Regression, Lasso, and Elastic Net. They are more advanced than simple linear regression as they reduce the variance of the model. In these three models, weights are assigned to features. Ridge regression and Lasso take penalty in sum of absolute values and sum of squared absolute values of weights respectively. Elastic net is a combination of Ridge and Lasso.

We build the three models using all features and compare their goodness of fit, and then will examine their prediction power on testing data.

\pagebreak

## Results

###Exploratory Analysis

We assume that the government aims to match its budget and its income by adjusting the mill rates
1. data processing
```{r, warning=F, message=F}
## Aggregated Data
assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))[,-1]
#assessment_2020 <- read.csv(here("data","assessment_2020.csv"))[,-1]
#assessment_2020.full <- read_excel(here("data/Assessment", "2020 Only.xlsx"))
```

We calculated the total tax income of each municipality by summing the product of assessment total and mill rate of all tax class. Then we plotted the computed tax income vs municipal budget into line-chart to visualize their relationship. We chose Vancoouver, Richmond and Burnaby as representative because they are the three largest municipalities in Metro Vancouver. The plots showed that the computed tax amount was positively and approximatly linearly related to municipal budget through time.

@@

###Linear Model

**Table 2: Result from Model Used**

Model | Mutiple R-Squared | Adjusted R_Squared | MSE | PMSE 
------|--------|--------|--------|--------
      OLR full | 0.8874| 0.8707| 1.9843| 2.5902
      OLR reduced| 0.8874| 0.8721| 1.9845| 2.5237
      Ridge|  0.8868| NA| 1.9896| 2.5675
      LASSO| 0.8873| NA| 1.9855| 2.5280
      Elastic Net| 0.8625| NA| 2.2366| 2.5486

Above is the goodness of fit of five models. Except for the Elastic Net model which has higher MSE and PMSE, there are no significant differences between these models. We have not used cross-validation to reduce model overfitting.
Although the results are relatively similar and have high R_squared across all models, we believe after using cross-validation there will be a clearer difference between the goodness of fit of each model.

For a dtailed look of how the models were fitted and evaluated, please refer to the appendix.

## Conclusion

From our exploratory data analysis, we can see that assessment total, land total, tax code, year and municipalities are all significant in our model. The client suggested that we can transform numerical factors such as assessment total into percentage change. This data transformation does not perform as well as we expected, which only yields an R^2 of around 0.2 across all of our fitted models. We believe the method that the client has suggested might leave out some important information about the housing market in each municipality. We have also found that when comparing the correlation between mill rate against features in our models, Vancouver is an outlier. We would like to further investigate Vancouver as a special case. We are also interested in the effect of different tax classes in predicting mill rate. Further analysis and prediction will be performed based on our hypothesis.

All fitted linear model were able to make accurate prediction based on means squared error, but there is still a very high chance that our model is overfitted. For our next report, we are going to use cross-validation to reduce the effect of overfitting.

## References

Links to source of data:

- Schedule 706 (https://www2.gov.bc.ca/gov/content/governments/local-governments/facts-framework/statistics/statistics)

Code repository:

- Data Cleaning (https://github.com/STAT450-550/RealEstate/blob/450/src/Data_Cleaning.Rmd)

- Exploratory Data Analysis and Model Fitting (https://github.com/STAT450-550/RealEstate/blob/450/src/EDA%26mode_fitting.Rmd)

## Appendix

### Missing Value:

There are 1801 missing values in mill rate(TaxClassTaxRate). We decided to impute these missing values
Based on client information, all properties in the same region, classcode, and year should have a unique class rate

- For entries with mill rate, we aggregated them into groups by region + classcode + year.

- For entries without mill rate, we found the group they belong to and assign them mill rate in that group.

###Here is some exceptions found:

Some groups' mill rate is not unique:

- In Delta, properties in different neighbourhoods have slightly different mill rates. Since the variance is not significant, we take the mean as the overall mill rate in groups.

- In Vancouver, 2019, Class 01, one property's mill rate is different from others. It is regarded as an outlier.

- In Burnaby, 2019, Class 06, six properties' mill rate are different from others. They are regarded as outliers.

- In Langley, 2019, Class 06, mill rate is different between assessment type. After talking to the client, the mill rate for assessment type "land" is regarded as the overall mill rate in that group.

- In some groups, all entries' mill rates are missing. Entries in these groups are removed. Here is the list of the groups:

**Table 3: Data with missing mill rate**

Year  | Region | Class |Number of Properties
------|--------|-------|----------
2016 | Belcarra | 06 | 9 
2016 | Lions Bay | 01 | 40 
2016 | Lions Bay | 06 | 25 
2016 | Maple Ridge Rural | 05 | 36 
2017 | Belcarra | 06 | 9 
2017 | Lions Bay | 01 | 39 
2017 | Lions Bay | 06 | 24 
2017 | Maple Ridge Rural | 05 | 36 
2018 | Maple Ridge Rural | 05 | 36 
2019 | Maple Ridge Rural | 05 | 38
